---
title: "hw3 data visualization"
author: "Yutian Luo"
date: "10/6/2020"
output: github_document
---

```{r setup, include=FALSE}
# # embed the plots
# knitr::opts_chunk$set(
#   fig.width = 5,
#   fig.asp = 9,
#   out.width = "90%"
# )
```

```{r}
# install.packages("devtools")
devtools::install_github("p8105/p8105.datasets")

library(p8105.datasets)
library(tidyverse)
library(dplyr)
library(ggridges)
library(hexbin)
data("instacart")
```


# Problem 1


The goal is to do some exploration of this dataset. 

To that end, write a short description of the dataset, 
noting the size and structure of the data, 
describing some key variables, 
and giving illstrative examples of observations





How many aisles are there, and which aisles are the most items ordered from?

Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 1000) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```




Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog fodd care", "packaged vegetables fruits")) %>% 
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(aisle, rank) %>% 
  knitr::kable()

```


Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r}
instacart %>% 
  filter(product_name %in% c("Pink Lady Apples","Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour = mean(order_hour_of_day)) %>% 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  ) %>% 
  knitr::kable()
```









## Problem 2

In this spreadsheet, variables activity.* are the activity counts for each minute of a 24-hour day starting at midnight.

Q:
Load, tidy, and otherwise wrangle the data. 
Your final dataset should include all originally observed variables and values; 
have useful variable names; 
include a weekday vs weekend variable; 
and encode data with reasonable variable classes. 

```{r}
accel_data = 
  read_csv(file = "./data/accel_data.csv") %>% 
  # pivot to long format
  pivot_longer(
    activity.1 : activity.1440,
    names_to = "minute",
    values_to = "activity_count"
  ) %>% 
  mutate(
    weekday_vs_weekend = 
      ifelse(day == 'Friday' | day == 'Saturday', "weekend", "weekday"),
    weekday_vs_weekend = as.factor(weekday_vs_weekend),
    day = as.factor(day)
  ) %>%
  separate(minute, into = c("acti","minute"), sep = 9, remove = TRUE) %>% 
  mutate(
    minute = as.integer(minute)
  ) %>% 
  select(-acti) 
  
# take a look at the tidy data
head(accel_data) %>% 
knitr::kable()
```

Q:

Describe the resulting dataset (e.g. what variables exist, how many observations, etc).

A:

The dataset contains all the recorded activity of the patient, given that record interval is 1 minute. It contains variables `r names(accel_data)`, and it has a total dimension of `r dim_desc(accel_data)`. The total observation is `r count(accel_data)`.


Q:

Traditional analyses of accelerometer data focus on the total activity over the day.Using your tidied dataset, aggregate accross minutes to create a total activity variable for each day, 
and create a table showing these totals. 

```{r}
# create a table showing total activity
accel_data_df1 =
  accel_data %>% 
  group_by(day_id) %>% 
  summarize(total_activity = sum(activity_count))

# show the table
accel_data_df1 %>% 
  knitr::kable()
```


Q:

Are any trends apparent?

```{r}
accel_data_df2 = 
  right_join(x = accel_data, y = accel_data_df1) %>% 
  mutate(
    day = forcats::fct_relevel(day, c("Friday", "Saturday", "Sunday", "Monday", "Tuesday", "Wednesday", "Thursday"))
  )

# detect any trend using line plot and smooth line plot

# preliminary:
# scatterplot: total activity by day_id
accel_data_df2 %>% 
  ggplot(mapping = aes(x = day_id, y = total_activity)) + 
  geom_point()

# smooth curve: total activity by day_id
accel_data_df2 %>% 
   ggplot(aes(x = day_id, total_activity)) + 
  geom_point() + 
  geom_smooth()

# boxplot: total activity by day
accel_data_df2 %>% 
  ggplot(aes(x = day, total_activity)) + 
  geom_boxplot()

# density: total activity by weekend
accel_data_df2 %>% 
  ggplot(aes(x = total_activity, fill = weekday_vs_weekend)) +      geom_density(alpha = 0.5, adjust = 0.9)

```


A:

According to the scatter smooth curve, there is a limited evidence of trend, which indicates that the activity increases as the observational days increase, and it decreases after the 18th day. It returns back to a higher value at 35th day.   

In relation to each weekdays, each day have different means and range. However, Satureday has the largest range for activity. Wednesday and Tuesday have a very little range. Friday is the most active day.

Observe the weekday and weekend's relation with daily acitivity, we can say that weekdays have a more stable activity, while weekend activity often varies a lot.

It should be noted, that the observational time is limited in 5 weeks. Some more significant observations can be made about the weekday weekends' relation to daily activity if the time period is longer.


Q:

Accelerometer data allows the inspection activity over the course of the day.
Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. 


```{r}
accel_data %>% 
  ggplot(aes(x = minute, activity_count, color = day)) + 
  geom_line() +
  scale_x_continuous(
    breaks = c(0, 240, 480, 720, 960, 1200, 1440),
    labels = c("0:00", "04:00", "08:00", "12:00", "16:00", "20:00", "24:00")
  ) 
```

Q:

Describe in words any patterns or conclusions you can make based on this graph.


A:

Wednesday has the lowest activity level in the coures of day. The activity level has to high points: around noon and around 8pm - 9pm. We can also see that the patient is very active in Friday nights.




# Problem 3

```{r}
library(p8105.datasets)
data("ny_noaa")
```


Q:

The goal is to do some exploration of this dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and indicating the extent to which missing data is an issue


```{r explore dataset}
# describe data
ny_noaa
summary(ny_noaa)

# describe NA: id and date both have no missing
anyNA(ny_noaa$id)       
anyNA(ny_noaa$date)

na_tmax = sum(is.na(ny_noaa$tmax))
na_tmin = sum(is.na(ny_noaa$tmin))
na_prcp = sum(is.na(ny_noaa$prcp))
na_snow = sum(is.na(ny_noaa$snow))
na_snwd = sum(is.na(ny_noaa$snwd))
na_total = sum(is.na(ny_noaa))
```

A:

The dataset have a size of `r ncol(ny_noaa)` * `r nrow(ny_noaa)`. It describes the weather record of the area from `r min(ny_noaa$date)` to `r max(ny_noaa$date)`. The variables that describes weather metrics include: `r names(ny_noaa)[3:7]`. 

Among these data of `r nrow(ny_noaa) * ncol(ny_noaa)`, there are `r na_total` missing values, which takes up to `r na_total / (nrow(ny_noaa) * ncol(ny_noaa)) * 100` % of the observations. Among these, `r na_tmax / na_total * 100` % are from tmax, `r na_tmin / na_total * 100` % are from tmin. 



Q:

Do some data cleaning. 
Create separate variables for year, month, and day. 
Ensure observations for temperature, precipitation, and snowfall are given in reasonable units. 

```{r data cleaning}
ny_noaa_tidy = 
ny_noaa %>% 
  mutate(
    date = as.character(date),
    tmax = as.numeric(tmax),
    tmin = as.numeric(tmin),
    prcp = prcp / 10,
    tmax = tmax / 10,
    tmin = tmin / 10
  ) %>% 
  separate(date, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(
    year = as.numeric(year),
    month = as.numeric(month),
    day = as.numeric(day)
  )

# most commonly observed snowfall
snow_count = count(ny_noaa, snow) %>% 
  mutate(
    snow_rank = min_rank(desc(n))
  ) %>% 
  arrange(snow_rank, by_group = FALSE)

```

Q:

For snowfall, what are the most commonly observed values? Why?

A:

0. Since most of the days in a year it is not snowing.


Q:

Make a two-panel plot showing the average max temperature in January and in July in each station across years. 

```{r}
ny_noaa_df1 = 
ny_noaa_tidy %>% 
  filter(month == 1 | month == 7) %>%
  group_by(id, month, year, day) %>% 
  summarize(
    avg_tmax = mean(tmax, na.rm = TRUE)
  ) %>% 
  drop_na()
  
ggplot(ny_noaa_df1, aes(x = year, avg_tmax), show_legend = FALSE) + 
  geom_point(aes(color = id))+
  geom_smooth(se = FALSE)+
  facet_grid(.~month) + 
  theme(legend.position = "none")
```


Q:

Is there any observable / interpretable structure? Any outliers?


A:

The average max temperature is higher in July, by 20 degress approximately. In January, the variations in tmax is greater than that in July. Also, the average tmax fluctuates a lot over the years from 1980 to 2010, this implies some more interesting hypothesis to be analyzed relating years with temperature change.


Q:

Make a two-panel plot showing 
(i) tmax vs tmin for the full dataset 
(note that a scatterplot may not be the best option);
(ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.

```{r tmax vs tmin for the full dataset }
ny_df2_plot =
ny_noaa_tidy %>%
  ggplot(aes(x = tmax, tmin)) +
  geom_hex()
```

```{r distribution of snowfall values}
ny_noaa_df3 = 
ny_noaa_tidy %>% 
  filter(snow > 0 & snow < 100) %>% 
  group_by(year) %>% 
  drop_na() 

ny_df3_plot = 
ny_noaa_tidy %>% 
  ggplot(aes(x = snow, year, group = year))+
  geom_density_ridges()

```

```{r combine 2 plots in a panel}
##(ny_df2_plot + ny_df3_plot)
```




